Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area.

### 2022-07-06 outputs/paper/paper.qmd
Updates

@@

@@

poll_2011_2020 %>%
  select(Year, Pollutant, January, February, March, April, May, June, July, August, September, October, 
         November, December) %>%
  slice(c(2, 34, 74, 91, 117, 147, 184, 223, 252, 308)) %>%
  kable(
    caption = "Extracting rows from the Pollution data from 2011 - 2020",
    col.names=c("Year", "Pollutant", "Jan", "Feb", "Mar", "Apr", "May",
                "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
    linesep = "", digits = 2, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)

@@

@@

dec_mar <- temperature_2002_2022 %>%
  filter(month_name == "Dec" | month_name == "Jan" | month_name == "Feb" | month_name == "Mar")

ggplot(dec_mar, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")

@@

@@

apr_jul <- temperature_2002_2022 %>%
  filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul")

ggplot(apr_jul, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")

@@

@@

aug_nov <- temperature_2002_2022 %>%
  filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")

ggplot(aug_nov, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")

@@

>>>>

Urgent action is needed and the science shows that it will make a tangible difference. The health impacts from climate change can be prevented by improving efforts both in societies and health systems in Canada. Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change.

@@

Table 1 shows the rows of the Temperature data from 2002 to 2022. Variable "Year" indicates the year, variable "Month" indicates the month, and variable "Day" indicates the day. Variable "Max Temp" represents the maximum temperature on the day with the following month and year. Variable "Min Temp" represents the minimum temperature on the day with the following month and year. Variable "Mean Temp" represents the mean temperature on the day with the following month and year.

@@

pollution %>%
  select(city, pollutant, mean_pollution, day, month, year) %>%
  slice(c(100, 1000, 1200, 1500, 1800, 2000, 2500, 2800, 3600, 5000, 6000, 7000, 10000, 11000, 15000)) %>%
  kable(
    caption = "Extracting rows from the Pollution data from 2002 - 2020",
    col.names=c("City", "Pollutant", "Mean Pollution", "Day", "Month", "Year"),
    linesep = "", digits = 2, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)

@@

Table 2 shows the rows of the Pollution data from 2002 to 2020. Variable "City" indicates the city from which the data was collected. Variable "Pollutant" indicates the pollutant and includes 5 types: CO, NO, NO2, O3, and SO2. Variable "Mean Pollution" represents the mean pollution levels of the following pollutant on the following day of the month and year. Variable "Day" indicates the day, variable "Month" indicates the month, and variable "Year" indicates the year. Variable "Max Temp" represents the maximum temperature on the day with the following month and year.

@@

temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| 
               month_name == "Feb" | month_name == "Mar") %>% ggplot(
              aes(x=year, y=max_temp, color=month_name)) + 
  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")

@@

Figure 1 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from December to March.

@@

temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | 
                            month_name == "Jun" | month_name == "Jul") %>% ggplot(
                              aes(x=year, y=max_temp, color=month_name)) + 
  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")

@@

Figure 2 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from April to July.

@@

temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | 
                            month_name == "Oct" | month_name == "Nov") %>% ggplot(
                              aes(x=year, y=max_temp, color=month_name)) + 
  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")

@@

Figure 3 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from August to November.

## Visualizing Pollution from 2002-2020

#### Visualizing Pollution by Year

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Year", fig.width=9, fig.height=3.2}

pollution %>% filter(pollutant != "CO") %>% ggplot(
  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
```

Figure 4 shows an ... pattern for the mean pollution from year 2002 to 2020 for NO, NO2, O3, and SO2 by year.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Year", fig.width=9, fig.height=3.2}

pollution %>% filter(pollutant == "CO") %>% ggplot(
  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")

Figure 5 shows an ... pattern for the mean pollution from year 2002 to 2020 for CO by year.

#### Visualizing Pollution by Month

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Month", fig.width=9, fig.height=3.2}

pollution %>% filter(pollutant != "CO") %>% ggplot(
  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)")

Figure 6 shows an ... pattern for the mean pollution from year 2002 to 2020 for NO, NO2, O3, and SO2 by month.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month", fig.width=9, fig.height=3.2}

pollution %>% filter(pollutant == "CO") %>% ggplot(
  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")

Figure 7 shows an ... pattern for the mean pollution from year 2002 to 2020 for CO by month.

## Visualizing Pollution and Temperature from 2002-2020

# Load pollution and temperature data

pollu_temp <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_temperature_pollution.csv")
```

#### Visualizing Mean Pollution by Temperature, Pollutant and Month

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}

pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
  labs(color="Month") + xlab("Maximum Temperature") + 
  ylab("Mean Pollution")
```

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}

pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
  labs(color="Month") + xlab("Maximum Temperature") + 
  ylab("Mean Pollution")

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}

pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
  (month_name == "Apr" | month_name == "May" | month_name == "Jun" | 
  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, 
  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
  labs(color="Month") + xlab("Maximum Temperature") + 
  ylab("Mean Pollution")

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}

pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & (month_name == "Apr" |
   month_name == "May" | month_name == "Jun" | month_name == "Jul")) %>% ggplot(aes(max_temp,
   mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
  labs(color="Month") + xlab("Maximum Temperature") + 
  ylab("Mean Pollution")

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}

pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & (month_name == "Aug" | 
    month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp,
    mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
    labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}

pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
  (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>%
  ggplot(aes(max_temp, mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~
  pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")

In the above plot...
